---
title: '`ggplot2`Practice'
author: "Shannon Ellis"
date: "9/11/2018"
output: html_document
---

The goal of this document is to help get more comfortable plotting data using the `ggplot2` package.

### `ggplot2` review

Very briefly, the general format for making a plot using `ggplot2` is as follows:

```
ggplot(df, aes())
 + geom()
```

In this:

* `df` : represents the dataframe you want to plot
* `aes` : includes the mappings you'll use to generate teh plot
* `geom` : represents what `geom` you want to use (or, in other words, what *type* of plot you want to make)

For more detailed information about making plots using `ggplot2`, refer back to the Data Visualization Leanpub course.

### Install & Load `ggplot2`

To complete these exercises, you'll first have to make sure `ggplot2` is installed and loaded in.

```{r load-package}
## Install and load the ggplot2 package
install.packages("ggplot2")
install.packages("dplyr")
library(ggplot2)
library(dplyr)
```

### The Datasets

In these exercies we'll be using two datasets: `mtcars` and `msleep`.

You'll want to be sure you know what variables are available in each dataset. As before, use the functions `glimpse()` and `View()` to get reacquainted with both the `mtcars` and the `msleep` dataset. 

```{r view-data}
## Get acquainted with the dataset
glimpse(msleep)
glimpse(mtcars)
View(msleep)
View(mtcars)
```

### Histogram

Histograms are helpful for visualizing the *distribution* of a single numeric variable.

Choose a numeric (continuous) variable from the `mtcars` dataset and visualize the distribution of that variable using a histogram.

```{r histogram-mtcars}
## Include code for a histogram 
## for a single numeric variable in mtcars
ggplot(mtcars, aes(disp)) +
  geom_histogram()
```

Do the same for the `msleep` dataset. Choose a numeric (continuous) variable from the `msleep` dataset and visualize the distribution of that variable using a histogram.

```{r histogram-msleep}
## Include code for a histogram 
## for a single numeric variable in msleep
ggplot(msleep, aes(sleep_rem)) +
  geom_histogram()
```

### Scatterplot

Scatterplots are helpful for visualizing the relationship between two numeric variables. 

Choose two numeric (continuous) variables from the `mtcars` dataset and visualize the relationship between the two variables using a scatterplot

```{r scatterplot-mtcars}
## Include code for a scatterplot 
## for two numeric variables in mtcars
ggplot(mtcars, aes(mpg, cyl)) +
  geom_point()
```

Now, do the same for two numeric (continuous) variables from the `msleep` dataset. Vsualize the relationship between the two variables using a scatterplot.

```{r scatterplot-msleep}
## Include code for a scatterplot 
## for two numeric variables in msleep
ggplot(msleep, aes(sleep_total, sleep_rem)) +
  geom_point()
```


### Bar Charts

So far we've only been working with numeric variables, but it's important to be able to visualize categorical (factors) type data as well.

Bar charts are helpful for visualizing the number of observations in each level of a categorical variable.

Choose a categorical (factor) variable from the `mtcars` dataset and visualize the number within each level of the variable using a bar chart. 

(Note that sometimes categorical variables can look like a numeric variable, such as the number of cylinders a car has. However, it's a factor because there are only a few different options for how many cylinders a car can have.)

```{r bar-chart-mtcars}
## Include code for a barchart 
## for a categorical variable in mtcars
mtcars$vs <- as.factor(mtcars$vs)
ggplot(mtcars, aes(vs)) +
  geom_bar()
```

Do the same for the `msleep` dataset. Choose a categorical (factor) variable from the `mslee[` dataset and visualize the number within each level of the variable using a bar chart. 

```{r bar-chart-msleep}
## Include code for a barchart 
## for a categorical variable in sleep
msleep$vore <- as.factor(msleep$vore)
ggplot(msleep, aes(vore)) +
  geom_bar()
```

### Piping & `ggplot2`

Now, let's combine this with what you did in the piping practice lesson!

First, copy the code from the following two chunks (`pipe-carnivores` and `pipe-big-cars`) from the piping practice.

After adding the code from the piping exercise, create a scatterplot taking the output from the piping_practice code and piping it into the code that will generate a `ggplot2` scatterplot.

Use the same numeric variables you used above above for your scatterplots.

Compare the output from your scatterplots above to these plots. What has changed?

```{r pipe-carnivores}
## copy code from piping_practice
## pipe this code into code that will create
## a ggplot2 scatterplot
msleep %>%
  filter(order == "Carnivora") %>%
  arrange(sleep_rem, sleep_total) %>%
  ggplot(aes(sleep_total, sleep_rem)) +
  geom_point()
```


```{r pipe-big-cars}
## copy code from piping_practice
## pipe this code into code that will create
## a ggplot2 scatterplot
mtcars %>%
  filter(cyl == 4) %>%
  arrange(wt) %>%
  ggplot(aes(mpg, cyl)) +
  geom_point()
```
